---
title: "Data 606 Presentation"
author: "Natalie Mollaghan"
date: "11/30/2017"
output: html_document
---

#Data Preperation

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(plyr)
library(psych)
```

Load data into R

```{r}
refugeedata <- read.csv("https://raw.githubusercontent.com/ntlrs/data606finalproject/master/UN%20Data.csv", header = TRUE, stringsAsFactors = FALSE)
```

Clean Data
```{r}
refugee<-refugeedata[c(1:3)]
names(refugee)[1] <- "CoO"
names(refugee)[2] <- "Year"
names(refugee)[3] <- "Refugees"
refugee$CoO <- as.factor(refugee$CoO)
head(refugee)
```

###Research Question:
Which countries have the most had biggest increase in refugees since 1975? 

###Cases:
The UN has data on Refugees migration from 1975-2016

###Type of Study:
This is an observational study looking at data from 1975 to 2016 for all countries that have been forced to leave their country or territory of origin.

###Data Source:
The UN data on refugees can be found here: http://data.un.org/Data.aspx?d=UNHCR&f=indID%3AType-Ref#UNHCR

###Response:
The response variable is a numerical value that represents the number of refugees.

###Explanatory:
The explanatory variable will be calculated from the years and the number of refugees. 

###Relevant Summary Statistics
```{r}
summary(refugee)
```

```{r}
refugeeiraq <- subset(refugee, CoO == "Iraq")
plot(refugeeiraq$Year, refugeeiraq$Refugees, type = "l")
```
```{r}
sum(refugeeiraq$Refugees)
```
```{r}
describe(refugee$Refugees)
```



